Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-14 write-ups/PLDI/paper.tex
Shortening and polishing.

Because workgroups of a cooperative kernel are fairly scheduled, a
global barrier primitive can be provided.  We specify two such primitives: $\globalbarrier$
and $\resizingglobalbarrier$.

@@

Semantically, when all threads have reached $\resizingglobalbarrier$,
the number of active workgroups, $M$, is atomically set to a new value, $M'$ say, with $M' > 0$.
If $M' = M$ then the active workgroups remain unchanged and proceed to
execute after the barrier.  If $M' < M$, workgroups $[M', M-1]$ are
killed.  If $M' > M$ then $M'-M$ new workgroups join the computation,
as if they were forked from workgroup 0.  In particular, the
$\transmit$-annotated local state of thread 0 in workgroup 0 is
transmitted to the threads of the new workgroups.  The change in
number of active workgroups from $M$ to $M'$ is atomic.

@@

The enclosing $\globalbarrier$ calls ensure that the change in number
of active workgroups from $M$ to $M'$ occurs entirely within the
resizing barrier, so that from a programmer's perspective, $M$ changes atomically.  The middle $\globalbarrier$ is required to ensure that forking occurs before killing, so that while the barrier may kill a workgroup with id $i$, it should not also fork a new workgroup with id $i$.  That is, workgroups $[0, \mathit{min}(M, M') - 1]$ should be left intact.

Because $\resizingglobalbarrier$ can be implemented in the above
manner, we do not regard it conceptually as a primitive of the
cooperative kernels programming model.  However, we discuss in
\mysec\ref{sec:resizingbarrier} how $\resizingglobalbarrier$ can be
implemented significantly more efficiently if it is treated as a
primitive operation that interacts directly with the scheduler.

>>>>

Workgroups of a \emph{cooperative} kernel \emph{are} fairly scheduled, so a
global barrier primitive can be provided, and we specify two variants: $\globalbarrier$
and $\resizingglobalbarrier$.

@@

When all threads have reached $\resizingglobalbarrier$,
the number of active workgroups, $M$, is atomically set to a new value, $M'$ say, with $0 < M' \leq N$.
If $M' = M$ then the active workgroups remain unchanged.  If $M' < M$, workgroups $[M', M-1]$ are
killed.  If $M' > M$ then $M'-M$ new workgroups join the computation after the barrier,
as if they were forked from workgroup 0.  In particular, the
$\transmit$-annotated local state of thread 0 in workgroup 0 is
transmitted to the threads of the new workgroups.

@@

The enclosing $\globalbarrier$ calls ensure that the change in number
of active workgroups from $M$ to $M'$ occurs entirely within the
resizing barrier, so that from a programmer's perspective, $M$ changes atomically.  The middle $\globalbarrier$ ensures that forking occurs before killing, so that while the barrier may kill a workgroup with id $i$, it should not also fork a new workgroup with id $i$.  That is, workgroups $[0, \textrm{min}(M, M') - 1]$ should be left intact.

Because $\resizingglobalbarrier$ can be implemented in the above
manner, we do not regard it \emph{conceptually} as a primitive of our
programming model.  However, we discuss in
\mysec\ref{sec:resizingbarrier} how $\resizingglobalbarrier$ can be
implemented significantly more efficiently if it is treated as a
primitive operation that interacts directly with the scheduler.
